### 2014-01-19 proposal/proposal.tex
Skeleton for proposal and .gitignore.

\documentclass[10pt]{article}

% =========================================================================
% document style changes
% =========================================================================

\usepackage{amsmath}                    % AMS math packages
\usepackage{amssymb}                    %
\usepackage[]{graphpap}
\usepackage[T1]{fontenc}                % for \mathrm{}
\usepackage{courier}                    % for \texttt{}
\usepackage{bbm}                        % for \mathbbm{1} (indicator function)
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

%\setlength{\parskip}{\baselineskip}     % skip line following paragraphs
\setlength{\parskip}{0.0in}
\setlength{\parindent}{0.3in}             % Control margins and amount of text
\setlength{\topmargin}{-0.8in}
\setlength{\oddsidemargin}{-.15in}       % changed from {-.15in}
\setlength{\textheight}{9.5in}
\setlength{\textwidth}{6.8in}
\pagestyle{empty}                       % No page numbers

\newcommand{\spc}{\vspace{0.25in}}      % Shortcut commands
\newcommand{\ds}{\displaystyle}         %\newcommand{\ds}[1]{\displaystyle{#1}}
\newcommand{\ra}{\rightarrow}

\begin{document}                        % This is where the document begins

{\LARGE\bf
\begin{tabbing}
\hspace{2.8in} \= \hspace{1.3in} \= \hspace{1.2in} \= \\

%=========    Heading   ==================================================
CS 399 \> PA4 \> Bill Chickering (bchick)\\
\normalsize Jan 18, 2014 \> \> Jamie Irvine (jirvine)
% =========================================================================
\end{tabbing}
}
\vspace{.4in}

The abstract goes here. . . . 
Blah blah blah.  Blah blah blah.  Blah blah blah.  Blah blah blah.  Blah blah
blah.  Blah blah blah.  Blah blah blah.  Blah blah blah.  Blah blah blah.  Blah
blah blah.  Blah blah blah.  Blah blah blah.  Blah blah Blah blah blah.  Blah
blah blah.  Blah blah blah.  Blah blah blah.  Blah blah Blah blah blah.  Blah
blah blah.  Blah blah blah.  Blah blah blah.  Blah blah Blah blah blah.  Blah
blah blah.  Blah blah blah.  Blah blah blah.  Blah blah Blah blah blah.  Blah
blah blah.  Blah blah blah.  Blah blah blah.  Blah blah Blah blah blah.  Blah
blah blah.  Blah blah blah.  Blah blah blah.  Blah blah blah.  Blah blah blah.
Blah blah blah.  Blah blah blah.  blah.  Blah blah blah.  Blah blah blah.  Blah
blah blah.  blah.  Blah blah blah.  Blah blah blah.  Blah blah blah.  blah.
Blah blah blah.  Blah blah blah.  Blah blah blah.  blah.  Blah blah blah.  Blah
blah blah.  Blah blah blah.  blah.  Blah blah blah.  Blah blah blah.  Blah blah
blah.

\section*{Vision}

\section*{Implementation Ideas}

\end{document}