### 2014-02-17 proposal/new_proposal.tex
Added new proposal

\documentclass[10pt]{article}

% =========================================================================
% document style changes
% =========================================================================

\usepackage{amsmath}                    % AMS math packages
\usepackage{amssymb}                    %
\usepackage[]{graphpap}
\usepackage[T1]{fontenc}                % for \mathrm{}
\usepackage{courier}                    % for \texttt{}
\usepackage{bbm}                        % for \mathbbm{1} (indicator function)
    \usepackage{booktabs}
    \usepackage{graphicx}
    \usepackage{caption}
    \usepackage{subcaption}

    %\setlength{\parskip}{\baselineskip}     % skip line following paragraphs
    \setlength{\parskip}{0.0in}
    \setlength{\parindent}{0.3in}             % Control margins and amount of text
    \setlength{\topmargin}{-0.8in}
    \setlength{\oddsidemargin}{-.15in}       % changed from {-.15in}
    \setlength{\textheight}{9.5in}
    \setlength{\textwidth}{6.8in}
    \pagestyle{empty}                       % No page numbers

    \newcommand{\spc}{\vspace{0.25in}}      % Shortcut commands
    \newcommand{\ds}{\displaystyle}         %\newcommand{\ds}[1]{\displaystyle{#1}}
    \newcommand{\ra}{\rightarrow}

    \begin{document}                        % This is where the document begins

{\LARGE\bf
    \begin{tabbing}
    \hspace{2.8in} \= \hspace{1.3in} \= \hspace{1.2in} \= \\

        %=========    Heading   ==================================================
        CS 399 \> PA4 \> Bill Chickering (bchick)\\
        \normalsize Jan 18, 2014 \> \> Jamie Irvine (jirvine)
        % =========================================================================
        \end{tabbing}
} \vspace{.4in}

\section*{Abstract}
\emph{Will write this when the paper is done}

\section*{Introduction}
For any large retailer, it is valuable to know the relationship between
products they sell or could sell. This information can be used to build
recommendations for similar products, understand subsets of products that
appeal to certain users, or expand a catalogue to include new products that
will be liked by the existing user-base. Currently, many retailers already
leverage user traffic and content information to learn these relationships
within their catalogue. However, in spaces where competing retailers sell
many of the same products, such as department stores or movie vendors, more
information can be gleaned by examining the inter-product relationships that
competing retailers have built for their own catalogues.

In this paper, we focus on department stores. This space is prime for our
analysis since there are a number of competing stores that have overlapping
catalogues. We model the relationships between products with a
\emph{recommendation graph}, which is directed graph between the products in a
catalogue. A directed edge from product A to product B means that if a user
likes product A, he/she is likely to like product B. First we construct the
recommendation graphs of different department stores using the
recommendations given for each product on their websites. With these
original graphs, we aim to add meaningful edges within one retailers graph
(representing a new recommendation) and between different retailer graphs
(representing a candidate to add to a catalogue).

We compare three approaches to extending the recommendation graphs. The
first approach is a baseline that uses a content-driven similarity function
to add edges between items. It ignores any information from the original
recommendation graphs. The second approach focuses exclusively on the
recommendation graphs. This method uses an identity function to determine
which items in different graphs are the same and thus links the graphs
together. Using a random walk algorithm, we determine which unlinked items
should be linked. The third approach is uses a combination of the first two
to leverage both kinds of information.

\section*{Data}
We collect the data by scraping the websites of major departments stores for
products with their descriptions, prices, photos, etc. From each product
page, we also gather the website's top recommended items for that product.
For this paper, we scraped the catalogues of Macy's, Bloomingdale's, etc.
Each of these websites requires their own scraper, but the same data is
accessible from all of them, including at least four recommended products
for each product in the catalogue. This allows us to build a directed
recommendation graph for each department store.

It is worth noting that in reality, a company would have more data about
their own products than what they post on their website. However, the
recommendation graph is still quite informative and is accessible to anyone.
It is realistic that a company could gather graphs like this from each of
their competitors, as we have here.

\section*{Methods}
\subsection*{Method 1}
create similarity scores between items. If items are similar, we add edges
in both directions to the recommendation graph. To determine the threshold
of how high the score must be before adding an edge, we calculate the
average similarity between items that are connected in the original graph.
These edges can connect two items in the same catalogue or items across
catalogues.

\emph{More about the specifics of the function here}

\subsection*{Method 2}
The second method uses the recommendation graphs themselves to determine new
recommendation edges. Since there is a significant overlap of products
between department stores (x\% between Macy's and Bloomingdales for example),
        recognizing which products are the same connects the otherwise disjoint
        graphs of different companies. To identify which products are identical, we
        construct an identity function. \emph{This function works like this and was
            tested like this, so we know it works.}

            With connected graphs, we can apply a random walk algorithm to determine
            which new edges would be most appropriate to add. In the random walk, each
            product node is a source of a 'walker', which has equal probability of
            moving down each outgoing edge. We add edges based on the probability that a
            walker goes from one product to another. Since the bench-mark for an
            appropriate threshold is less obvious in this method, we use the top X
            scores, the number of edges above the threshold found in method 1. While
            this is somewhat arbitrary, it allows for a fair comparison between the
            methods.

            \emph{More about the specifics of the random walk algorithm here}

            \subsection*{Method 3}
            The final method leverages both content-similarity and the structure of the
            recommendation graphs. First, we combine the recommendation graphs using an
            identity function and run a random-walk algorithm, as in method 2. This
            determines how 'far,' in a sense, any product is from any other product. To
            determine if there should be an edge added from product A to product B, we
            use the content-similarity function of method 1 to compare A to every other
            product in the catalogue and sum these scores up with the weights from the
            random-walk originating at product B. Thus content-similarity from product A
            to a product that is near product B in the recommendation graph increases
            the likelihood that product A should be connected to product B. \emph{An equation
                would probably be good in here}. For this method, we also add the top X
                edges based on the threshold in method 1.

                \emph{More about the specifics of the implementation here}

                \section*{Evaluation}
                Empirically evaluating the methods is somewhat difficult with the data we
                have access to. With all of the user-traffic data, one could see if an added
                recommendation edge connects two products that are in fact co-clicked on
                co-purchased. Without this, the question of whether one product should be
                recommended for another can be answered with more subjective human
                judgement.

                For this experiment, we collect the top edges to be added for each of the
                three methods. We show a subset of them to a number of people who rate them
                in exchange for small sexual favors. The tester is asked 'How true is the
                following statement: If someone likes product 1, they will probably like
                product 2', followed by the pictures and descriptions of the two items of
                the directed edge. The tester answers on a 5 point scale and the results are
                aggregated. Each tester is shown edges from each of three methods, to
                decrease bias from a biased tester.

                \emph{More about how many testers we use}

                \section*{Results}
                \emph{We'll see}

                \section*{Conclusion}
                \emph{Will write this when the paper is done}

                \end{document}