### 2014-03-25 paper/paper.tex
Finished linear techniques section. Added graph. Going to bed

\documentclass[11pt]{article}

% =========================================================================
% document style changes
% =========================================================================

\usepackage{amsmath}                    % AMS math packages
\usepackage{amssymb}                    %
\usepackage[]{graphpap}
\usepackage[T1]{fontenc}                % for \mathrm{}
\usepackage{courier}                    % for \texttt{}
\usepackage{bbm}                        % for \mathbbm{1} (indicator function)
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage[font={small}]{caption}
\usepackage{subcaption}

%\setlength{\parskip}{\baselineskip}     % skip line following paragraphs
%\pagestyle{empty}                       % No page numbers
\setlength{\topmargin}{-.5in}
\setlength{\textheight}{9in}
\setlength{\oddsidemargin}{.125in}
\setlength{\textwidth}{6.25in}

\newcommand{\spc}{\vspace{0.25in}}      % Shortcut commands
\newcommand{\ds}{\displaystyle}         %\newcommand{\ds}[1]{\displaystyle{#1}}
\newcommand{\ra}{\rightarrow}
\newcommand{\Cov}{\mathrm{Cov}}
\DeclareMathOperator*{\argmax}{arg\!\max}
\DeclareMathOperator*{\argmin}{arg\!\min}

\begin{document}                        % This is where the document begins

\title{Improving Item-Item Similarity Estimation during Collaborative Filtering}
\author{Bill Chickering and Jamie Irvine\\
CS 399 with Anand Rajaraman\\
Stanford University}
\renewcommand{\today}{March 25, 2014}
\maketitle

\section*{Abstract}
\emph{Will write this when the paper is done}

\section*{Introduction}
Determining the similarity of two users or items is valuable for a number of
data-mining applications including recommender systems, product assortment and
differentiation, as well as branding. Collaborative filtering is often used to
exploit user-item rating data to make such estimates. This is typically done by
computing the Pearson correlation or cosine similarity of vector representations
of the users or items \cite{Su2009}. Given sufficient user-item rating data,
this approach can be very effective. However, when the rating data is sparse for
one or both users/items, these metrics can yield dramatic discrepancies compared
to values computed with sufficient data at a later point in time. Moreover,
these early, low-confidence measurements can be indistinguishable from later,
high-confidence measurements.  In this report, we describe two novel techniques
for estimating the true similarity of two users or items given limited common
rating data by incorporating confidence into the measurement.

To keep the present work focused, we confine ourselves to memory-based
collaborative filtering (CF) in which we work with a logical user-item matrix.
Each element of the matrix corresponds to a rating given by a particular user to
a particular item. Further, we choose to limit our analysis to item-item
similarity but note that these techniques should be equally effective at
improving user-user similarity estimation. In this representation, an item is a
vector of the ratings it has received from all users. We now formalize the
notion of item-item similarity. Two alternative models for the generation of
user-item ratings are useful in arriving at two different definitions of the
{\em true} item-item similarity.

\section*{Similarity}
\subsection*{True Item-Item Similarity}

In the first model, we imagine users are randomly presented items to rate. Given
a sufficient period of time, all users will eventually rate all items, creating
item vectors that contain provided ratings from all users. The true similarity
in this model of {\em random} rating, denoted $TrueSimRand$, is then the vector
product of two complete item vectors. That is,
\begin{align}
TrueSimRand(A, B) = \frac{\sum\limits_{u\in U}
r_{u,A}r_{u,B}}{\sqrt{\sum\limits_{u\in U} r_{u,A}^2}
\sqrt{\sum\limits_{u\in U} r_{u,B}^2}},
\end{align}
where $A$ and $B$ are items, $U$ is the set of all users and $r_{u,I}$ is the 
rating given by user $u$ to item $I$.

In the second model, we imagine an unlimited supply of users, each of whom
chooses to rate a subset of items. The lack of a rating is now meaningful since
it indicates a lack of interest of a particular user for a particular item. This
is captured by the model through a default rating of zero\footnote{We address
the issue of rating bias and the precise meaning of a zero valued ratings later
in this report.}. At any point in time, we may estimate the true similarity in
this model of {\em preferential} rating, by computing the cosine similarity of
the two item vectors. Only in the limit of inifinite time does this estimate
converge to the true similarity, which we denote $TrueSimPref$.  That is,
\begin{align}
TrueSimPref(A, B) = lim_{t\to\infty}\frac{\sum\limits_{u\in U_{AB}}
r_{u,A}r_{u,B}}{\sqrt{\sum\limits_{u\in U_A} r_{u,A}^2}
\sqrt{\sum\limits_{u\in U_B} r_{u,B}^2}},
\end{align}
where $U_A$ is the set of users who have rated $A$, $U_B$ is the set of users
who have rated $B$, and $U_{AB}$ is the set of users who have rated both $A$ and
$B$.

\subsection*{Estimating Item-Item Similarity}

Of course, most systems will never acquire ratings from all users for all items.
Nor are we able to work in the limit of infinite time. We must therefore make
estimates of the true item-item similarity using only partial information. In
the case of $TrueSimPref$, the cosine similarity computed at any point in time
serves as a simple and obvious estimate. In using this estimate, we are treating
all missing ratings as intentionally omitted. When estimating $TrueSimRand$, the
choice of how to handle missing ratings is perhaps less obvious. Following the
convention of Breese et al. \cite{Breese1998}, we choose to use the Pearson
correlation\footnote{The Pearson correlation is traditionally characterized by
subtracting an expectation value from each dimension. This is only approximately
true in our case. We discuss the how we subtract rating biases later in this
report.} for this estimate, ignoring all ratings by users who have not rated
both items. To summarize, our naive estimates are
\begin{align}
PearSim(A, B) = \frac{\sum\limits_{u\in U_{AB}}
r_{u,A}r_{u,B}}{\sqrt{\sum\limits_{u\in U_{AB}} r_{u,A}^2}
\sqrt{\sum\limits_{u\in U_{AB}} r_{u,B}^2}}
\stackrel{?}{\approx} TrueSimRand(A, B)
\end{align}

\begin{align}
CosSim(A, B) = \frac{\sum\limits_{u\in U_{AB}}
r_{u,A}r_{u,B}}{\sqrt{\sum\limits_{u\in U_A} r_{u,A}^2}
\sqrt{\sum\limits_{u\in U_B} r_{u,B}^2}}
\stackrel{?}{\approx} TrueSimPref(A, B).
\end{align}

Both similarity estimates primarily leverage information from common users, that
is users who have rated both items. Because of this, they perform well with a
large number common users, but give unreliable, and dramatically varying,
results when the items have few common users. In an extreme case, if only one
user has rated item $A$ and item $B$, $PearsSim(A, B) = 1$ or $-1$. Not only is
this unlikely to be an accurate assessment of the true similarity of $A$ and
$B$, it also gives the most extreme results possible, without giving any
indication that this is a low-confidence calculation.

In the following sections, we present two approaches to better approximate the
true similarity given a limited number of common reviewers. Both approaches are
motivated by probabilistic models with distributions grounded in the data. The
first approach leverages the observed similarity score and the number of common
users to calculate the true similarity. The second approach leverages the
rating-vectors themselves.

\section*{Score-based Approach}
\subsection*{Probabilistic Model}

In our first approach, we model the probability distributions of true and 
observed similarity. The goal is to use the observed similarity score directly 
to approximate the true similarity. In this way, we can consider an abstact 
$TrueSim$ which ignores the details of $TrueSimRand$ or $TrueSimPref$. 
Similarly, we consider $Sim_n$ to be a generic observed similarity score when
the two itmes have $n$ common reviewers. Now, let $Y$ be a random variable 
representing the $TrueSim$ of a randomly chosen pair of items. Let $X_{n}$ be a
random variable representing the $Sim_{n}$ of a pair of items when $n$ common 
users are observed.

We model $Y$ as a Normal distribution 
\begin{align}
Y \sim N(\mu, \sigma_{s}^2)
\end{align}
where $\mu$ and $\sigma_{s}^2$ are the average similarity score and the variance 
of similarity scores of all pairs of items, respectively. Since $X_n$ represents
a noisy reading of the true similarity $Y$, we model $X_n | Y=y$ as a Gaussian
error around $y$:
\begin{align}
(X_n | Y=y) \sim N(y, \sigma_{err, n}^2)
\end{align}

Note that $\sigma_{err, n}^2$ represents how noisy $Sim_n$ is as an estimation of
$TrueSim$ and therefore should decrease as $n$ increases. The probability
distribution of $X_n$ can be visualized in Figure 1.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.7\textwidth]{twonormals.png}
	\caption{The blue curve represents $Y$, the distribution of true
    similarities between a random pair of items. The red curve represents $X|Y$,
    the observed similarity when a small number of common users exist.}
    \label{fig:two_normals}
\end{figure}

The problem of approximating the $TrueSim$ from $Sim_n$ can now be seen as
finding the value of $Y$ that most likely produced $X_n$. In other words we 
desire the Maximum Likelihood Estimate of $Y | X_n$:
\begin{align}
\hat{y} &= \argmax_yP(Y=y|X_n=x) 
\\&= \argmax_y\left[\frac{P(X_n=x | Y=y)P(Y=y)}{P(X_n=x)}\right]
\\&= \argmax_y\left[P(X_n=x|Y=y)P(Y=y)\right]
\\&= 
\argmax_y\left[\frac{1}{\sigma_{err,n}\sqrt{2\pi}}\exp{\left(\frac{-(x-y)^2}
{2\sigma_{err,n}^2}\right)}
\frac{1}{\sigma_{s}\sqrt{2\pi}}\exp{\left(\frac{-(y-\mu)^2}
{2\sigma_{s}^2}\right)}\right]
\\&= \argmin_y\left[\frac{(x-y)^2}{2\sigma_{err,n}^2} +
\frac{(y-\mu)^2}{2\sigma_{s}^2}\right]
\\&= \argmin_y\left[\left(\sigma_{s}^2+\sigma_{err,n}^2\right)y^2 - 
2\left(\sigma_{s}^2x+\sigma_{err,n}^2\mu\right)y\right]
\end{align}
We find the exact minimum by taking the derivative with respect to $y$ and
setting it to zero:
\begin{align}
\frac{d}{dy}\left[\left(\sigma_{s}^2+\sigma_{err,n}^2\right)y^2 - 
2\left(\sigma_{s}^2x+\sigma_{err,n}^2\mu\right)y\right]
&= 2\left(\sigma_{s}^2+\sigma_{err,n}^2\right)y - 
2\left(\sigma_{s}^2x+\sigma_{err,n}^2\mu\right) 
\\&= 0
\end{align}
This produces the following linear equation for $\hat{y}$ in terms of $x$:
\begin{align}
\hat{y} &= \frac{\sigma_{s}^2x+\sigma_{err,n}^2\mu}{\sigma_{s}^2+\sigma_{err,n}^2}
\end{align}
which can be rewritten as:
\begin{align}
\left(\hat{y} - \mu\right) &= \frac{\sigma_{s}^2}{\sigma_{s}^2+\sigma_{err,n}^2}
\left(x-\mu\right)
\end{align}

The model suggests that there is a linear correlation between $X_n$ ($Sim_n$)
and $Y$ ($TrueSim$). Moreover, the best predicted $TrueSim$ is a linear 
combination of the average true similarity and the observed similiarity. Since $\sigma_{err,n}^2$ is 
nonnegative, the slope is always less than or equal to one, meaning that the 
observed distance from $\mu$ (the average true similarity) overestimates the 
true distance. This is reasonable, since prior suggests that the true 
similarity is likely to be near $\mu$.\footnote{It is worth noting that despite
the resemblance to Slope One Predictors \cite{Lemire1998}, this is a fundamentally 
different technique. Here, we have a linear relationship between observed 
scores and true scores, whereas the Slope One model seeks a linear relationship 
between the ratings a user gives to one items versus another.}

This model requires a different parameter $\sigma_{err,n}^2$ for each $n$.
To gain a more generality, we model $\sigma_{err,n}^2$ as a function of $n$
instead of considering them independent of each other. Recall that 
$\sigma_{err,n}^2$ is the noisiness of the observed $Sim_n$ about $TrueSim$. 
Since the $TrueSim$ is the $Sim$ when all users that will rate the items 
have rated them, $Sim_n$ can be thought of as sampling $n$ users from the 
set used to calculate $TrueSim$. Although $Sim$ is likely more complicated 
than an average, the Central Limit Theorem motivates the intuition that the 
variance of $Sim$ would decrease as the sqaure root of $n$. Thus we model 
$\sigma_{err,n}^2$ as:
\begin{align}
\sigma_{err,n}^2 = \frac{\alpha}{\sqrt{n}},
\end{align}
yielding the single parameter model:
\begin{align}
\left(Y - \mu\right) = \frac{\sigma_{s}^2}{\sigma_{s}^2+\frac{\alpha}{\sqrt{n}}}
\left(X_n-\mu\right)
\end{align}

\subsection*{Technique}
While it may be impossible to get $TrueSim$ for any pair, we
can get close by calculating $Sim_n$ when a large number of users have rated
both items. Formally, we define $GoldSim$ = $Sim_N$ where $N$ is a number 
large enough that $GoldSim \approx TrueSim$.

With a measurable $GoldSim$ as our goal, we can apply supervised learning 
techniques. First, we find pairs of items that have at least $N$ common users 
and calculate their $GoldSim$. Next we withold some of the users to produce 
estimated $Sim_n$ for $n=1,...,N$. By doing this over many pairs of items, we 
produce pairs of ($Sim_n$, $TrueSim$) that can be used for training and testing.

Now, the parameters $\mu$, $\sigma_{s}^2$, and $\sigma_{err,n}^2$ can be 
approximated using the method of moments over the training data. The values of 
$\sigma_{err,n}^2$ are fit to the function $\frac{\alpha}{\sqrt{n}}$ and the $\alpha$ 
that minimizes the sum of squared error is computed.

The following graph compares the results of three linear fits, with varying
degrees of fidelity to the probabilistic model. The most faithful fit uses
$\mu$, $\sigma_{s}^2$ and $\alpha$ as calculated above to predict the $GoldSim$
using Eqn (18). The next fit uses the calculated values of $\sigma_{err,n}^2$ directly,
following Eqn (16) for predictions. The third fit simply uses a series of linear
regressions, one for each $n$. This model only assumes that for each $n$,
$Sim_n$ and $GoldSim$ are linearly correlated.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.7\textwidth]{LinearTechniques.png}
	\caption{Comparison of different linear models. For each number of reviewers
    in common $n$, the graphs represent the average squared error of predicted
    $GoldSim$ calculated from the observed $Sim_n$. As a baseline, the technique
    of guessing the average $GoldSim$ is included.}
    \label{fig:Linear}
\end{figure}

The linear regression technique trades the generality of the model for a better
fit on the training data. As the results show, the trade-off is worth it.

\section*{Ratings-based Approach}
\subsection*{Probabilistic Model}

We now construct a different probabilistic model based on the pairs of ratings
themselves. This model is readily exploited for making estimates of
$TrueSimRand$ and, to a lesser extent, $TrueSimPref$. In addition, this approach
has the added benefit of allowing us to infer implied predictions of
$TrueSimRand$ made by individual users each time they rate two items. This
latter feature allows us to define a new metric for users that we call {\em
user-predictivity}, which allows us to ascribe varying levels of confidence to
their contributions to a similarity estimate.

Let $Y$ be a random variable representing the $TrueSimRand$ of a randomly chosen
pair of items. As before, we model $Y$ as a Normal distribution
\begin{align}
Y \sim N(\mu_s, \sigma_s^2),
\end{align}
where $\mu_s$ and $\sigma_s^2$ are the average and variance of all true
similarity scores, respectively. Now, let $\vec{R} = \left(R_A~~R_B\right)$ be a
random two element vector representing the ratings given to items $A$ and $B$.
We model the conditional distribution of $\vec{R} | Y$ as a multivariate Normal
distribution
\begin{align}
(\vec{R} | Y=y) \sim N(\vec{\mu_r}, \mathbf{\Sigma_y}),
\end{align}
where the elements of $\vec{\mu_r}$ are the average ratings and
$\mathbf{\Sigma_y}$ is the $y$-dependent $2\times2$ covariance matrix.  Since
$y$ is the $TrueSimRand$ score, it can be thought of as the Pearson correlation
over the distribution of ratings $R_A | Y$ and $R_B | Y$. Assuming that the
variance of individual ratings $\sigma_r^2$ is independent of $y$, we have 
\begin{align}
y = \frac{\Cov{\left[R_A | Y=y , R_B | Y=y\right]}}{\sigma_r^2},
\end{align}
It therefore follows that if the ratings are indeed normally distributed 
then the covariance matrix is related to the true similarity $y$ by
\begin{align}
\mathbf{\Sigma_y} = \sigma_r^2 \cdot
\left[ \begin{array}{cc}  1 & y \\ y & 1 \end{array} \right].
\end{align}

Using this model, we may view the problem of estimating the similarity $Y$ of
items $A$ and $B$ that most likely produced ratings $R_A$ and $R_B$ as
that of finding the Maximum Likelihood Estimate of $Y | \vec{R}$:
\begin{align}
\hat{y} &= \argmax_yP(Y=y|\vec{R}=\vec{r}) 
\\&= \argmax_y\left[\frac{P(\vec{R}=\vec{r} |
Y=y)P(Y=y)}{P(\vec{R}=\vec{r})}\right]
\\&= \argmax_y\left[P(\vec{R}=\vec{r}|Y=y)P(Y=y)\right]
\\&=
\argmax_y\left[\frac{1}{2\pi\sqrt{\left|\mathbf{\Sigma_y}\right|}}\exp{\left(-\frac{1}{2}
(\vec{r} - \vec{\mu_r})^\mathrm{T}\mathbf{\Sigma_y}^{-1}(\vec{r} - \vec{\mu_r})\right)}
\frac{1}{\sigma_{s}\sqrt{2\pi}}\exp{\left(\frac{-(y-\mu_s)^2}
{2\sigma_{s}^2}\right)}\right]
\\&= \argmin_y\left[(\vec{r} -
\vec{\mu_r})^\mathrm{T}\mathbf{\Sigma_y}^{-1}(\vec{r} - \vec{\mu_r}) + \frac{(y -
\mu_s)^2}{\sigma_{s}^2} + \ln{\left|\mathbf{\Sigma_y}\right|} \right]
\\&=
\argmin_y\left[\vec{r^\prime}^\mathrm{T}\mathbf{\Sigma_y}^{-1}\vec{r^\prime}
 + \frac{(y - \mu_s)^2}{\sigma_{s}^2} + \ln{\left|\mathbf{\Sigma_y}\right|} \right]
\\&= \argmin_y\left[\frac{1}{\sigma_r^2}\frac{1}{1 -
y^2}\left[r_A^\prime(r_A^\prime - yr_B^\prime) + r_B^\prime(r_B^\prime -
yr_A^\prime)\right] + \frac{(y - \mu_s)^2}{\sigma_{s}^2} +
\ln{\left[\sigma_r^2(1 - y^2)\right]} \right],
\end{align}
where we used the fact that
\begin{align}
\mathbf{\Sigma_y}^{-1} = \frac{1}{\sigma_r^2}
\left[ \begin{array}{cc}  1 & y \\ y & 1 \end{array} \right]^{-1} =
\frac{1}{\sigma_r^2}\frac{1}{1 - y^2}\left[ \begin{array}{cc}  1 & -y \\ -y & 1
\end{array} \right].
\end{align}
and that
\begin{align}
\left|\mathbf{\Sigma}\right| = \sigma_r^2(1 - y^2)
\end{align}
along with the substitution $\vec{r^\prime} = (r_A^\prime~~r_B^\prime) = \vec{r}
- \vec{\mu_r}$. Next, we find the minimum by taking the derivative with respect
to $y$ and setting it to zero:
\begin{align}
\frac{d}{dy}\left[ \frac{1}{\sigma_r^2}\frac{1}{1 -
y^2}\left[r_A^\prime(r_A^\prime - yr_B^\prime) + r_B^\prime(r_B^\prime -
yr_A^\prime)\right] + \frac{(y - \mu_s)^2}{\sigma_{s}^2} +
\ln{\left[\sigma_r^2(1 - y^2)\right]} \right] &= 0 \\
-r_A^\prime r_B^\prime y^2 + ({r_A^\prime}^2 + {r_B^\prime}^2)y - r_A^\prime
r_B^\prime + \frac{\sigma_r^2}{\sigma_s^2}(1 - y^2)^2(y - \mu_s) - \sigma_r^2(1
- y^2)y &= 0.
\end{align}
This final expression implicitly yields the MLE of $y$ as a function of the
ratings $r_A$ and $r_B$. That is, by constraining the parameters $\mu_r$,
$\sigma_r$, $\mu_s$, and $\sigma_s$ to realistic values (e.g. $\sigma_r,
\sigma_s > 0$), this expression defines a surface that can be used to
efficiently estimate $y$ given a pair of ratings\footnote{Note that this
equation neglects the constraint that $-1 \leq y \leq 1$. As a consequence, some
combinations of ratings and parameters can yield MLEs for $y$ that exceed $1$ in
magnitude. This is easily fixed, however, by applying this additional
constraint.}.

\subsection*{User Predictivity}

As already mentioned, the estimator $PearSim$ is unable to say much about the
similarity of two items when only provided ratings from a single common user. In
this case, $PearSim$ yields $\pm1$ (or the indeterminate value $0/0$ in the case
that both ratings are zero) depending on whether the ratings have the same sign
or not. The MLE provided by the above model rectifies this by generating
similarity estimates in the range $[-1, +1]$ given a single pair of ratings. We
therefore consider each user who rates a pair of items as implicitly predicting
the similarity of the items. A key discovery of the present work is that the
accuracy of users' predictions in a training set is correlated with the accuracy
in a test set. We exploit this finding by ascribing a quantity we call {\em user
predictivity} to each user defined as
\begin{align}
p_u = e^{-\epsilon_u/\lambda},
\end{align}
where $\lambda$ is a parameter to be learned and
\begin{align}
\epsilon_u = \frac{1}{\left|I_u\right|}
\sum_{A,B \in I_u}{\left|ModelSimRand(A, B) - GoldSimRand(A, B)\right|}
\end{align}
with $I_u$ the set of item pairs rated by user $u$ such that $\epsilon_u$ is
the average absolute error of user $u$'s similarity predictions.

\section*{Experiment}

\section*{Results}

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.7\textwidth]{RatingDist_40.png}
    \includegraphics[width=0.7\textwidth]{RatingDist_80.png}
	\caption{Distribution of rating pairs by common reviewers for item pairs.
Top figure shows distribution for item pairs having at least 40 common
reviewers. Bottom figure shows distribution for item pairs having at least 80
common reviewers.}
    \label{fig:RatingDist}
\end{figure}

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=1.0\textwidth]{Histograms.png}
	\caption{Histograms of {\em TrueSimR} (top) and {\em TrueSimP} (bottom) for
training datasets containing at least 40 common reviewers (left) and 80 common
reviewers (right).}
    \label{fig:Histograms}
\end{figure}

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.9\textwidth]{MSE_SimR_40.png}
    \includegraphics[width=0.9\textwidth]{MSE_SimR_80.png}
	\caption{Mean square error {\em vs} number of common reviewers, $n$, for
estimates of {\em TrueSimR} for item pairs ultimately having at least 40 common
reviewers (top) and 80 common reviewers (bottom). The dashed curve represents
{\em SimR} computed using only the ratings from the first $n$ reviewers. The
flat red line indicate the MSE that results from always guessing the average
{\em TrueSimR} from the training set. The green curve is {\em RegSimR} and the
blue curve is {\em ModelSimR}. }
    \label{fig:MSE_SimR}
\end{figure}

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.9\textwidth]{MSE_SimP_80.png}
	\caption{Mean square error {\em vs} number of common reviewers, $n$, for
estimates of {\em TrueSimP} for item pairs ultimately having at least 80 common
reviewers. The dashed curve represents
{\em SimP} computed using only the ratings from the first $n$ reviewers. The
flat red line indicate the MSE that results from always guessing the average
{\em TrueSimP} from the training set. The green curve is {\em RegSimP} and the
blue curve is {\em ModelSimP}. }
    \label{fig:MSE_SimP}
\end{figure}

\section*{Conclusion}
\section*{Future Work}

\begin{thebibliography}{9}

\bibitem{Su2009}
    Xiaoyuan Su and Taghi M. Khoshgoftaar.
    ``A Survey of Collaborative Filtering Techniques.''
    \emph{Advances in Artificial Intelligence.} vol. 2009,
    Article ID 421425, 19 pages, 2009. doi:10.1155/2009/421425
\bibitem{Breese1998}
    Breese, John S., David Heckerman, and Carl Kadie, 
    ``Empirical analysis of predictive algorithms for collaborative filtering.'' 
    \emph{Proceedings of the Fourteenth conference on Uncertainty in artificial intelligence.}
    Morgan Kaufmann Publishers Inc., 1998.
\bibitem{Lemire1998}
    Daniel Lemire and Anna Maclachlan.
    ``Slope One Predictors for Online Rating-Based Collaborative Filtering.''
    \emph{SDM.} Vol. 5, 2005.

\end{thebibliography}

\end{document}