### 2014-03-26 paper/paper.tex
Make minor tweaks.

\documentclass[11pt]{article}

% =========================================================================
% document style changes
% =========================================================================

\usepackage{amsmath}                    % AMS math packages
\usepackage{amssymb}                    %
\usepackage[]{graphpap}
\usepackage[T1]{fontenc}                % for \mathrm{}
\usepackage{courier}                    % for \texttt{}
\usepackage{bbm}                        % for \mathbbm{1} (indicator function)
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage[font={small}]{caption}
\usepackage{subcaption}

%\setlength{\parskip}{\baselineskip}     % skip line following paragraphs
%\pagestyle{empty}                       % No page numbers
\setlength{\topmargin}{-.5in}
\setlength{\textheight}{9in}
\setlength{\oddsidemargin}{.125in}
\setlength{\textwidth}{6.25in}

\newcommand{\spc}{\vspace{0.25in}}      % Shortcut commands
\newcommand{\ds}{\displaystyle}         %\newcommand{\ds}[1]{\displaystyle{#1}}
\newcommand{\ra}{\rightarrow}
\newcommand{\Cov}{\mathrm{Cov}}
\DeclareMathOperator*{\argmax}{arg\!\max}
\DeclareMathOperator*{\argmin}{arg\!\min}

\begin{document}                        % This is where the document begins

\title{Improving Item-Item Similarity Estimation during Collaborative Filtering}
\author{Bill Chickering and Jamie Irvine\\
CS 399 with Anand Rajaraman\\
Stanford University}
\renewcommand{\today}{March 25, 2014}
\maketitle

\section*{Abstract}
\emph{The problem of inaccurate item-item similarity estimates due to a lack of
data is explored within the context of collaborative filtering. An off-line
experiment that evaluates the performance of similarity estimators is described.
Two novel item-item similarity estimators are described and analyzed. The first
is based on a model of noisy similarity score generation and proves effective at
estimating both future Pearson correlations and cosine similarities. The second
models noisy ratings as generated by similarity score dependent distributions.
Using the notion of user-predictivity, accurate estimates of future Pearson
correlation are made.}

\section*{Introduction}
Determining the similarity of two users or items is valuable for a number of
data-mining applications including recommender systems as well as product
assortment and differentiation. Collaborative filtering is often used to exploit
user-item rating data to make such estimates. This is typically done by
computing the Pearson correlation or cosine similarity of vector representations
of the users or items \cite{Su2009}. Given sufficient user-item rating data,
this approach can be very effective. However, when the rating data is sparse for
one or both users/items, these metrics can yield dramatic discrepancies compared
to values computed with sufficient data at a later point in time. Moreover,
these early, low-confidence measurements can be indistinguishable from later,
high-confidence measurements.  In this report, we describe two novel techniques
for estimating the true similarity of two users or items given limited common
rating data by incorporating confidence into the measurement.

To keep the present work focused, we confine ourselves to memory-based
collaborative filtering (CF) in which we work with a logical user-item matrix.
Each element of the matrix corresponds to a rating given by a particular user to
a particular item. Further, we choose to limit our analysis to item-item
similarity but note that these techniques should be equally effective at
improving user-user similarity estimation. In this representation, an item is a
vector of the ratings it has received from all users. We now formalize the
notion of item-item similarity. Two alternative models for the generation of
user-item ratings are useful in arriving at two different definitions of the
{\em true} item-item similarity.

\section*{True Item-Item Similarity}

In the first model, we imagine users are randomly presented items to rate. Given
a sufficient period of time, all users will eventually rate all items, creating
item vectors that contain provided ratings from all users. The true similarity
in this model of {\em random rating}, denoted $TrueSimRand$, is then the vector
product of two complete item vectors. That is,
\begin{align}
TrueSimRand(A, B) = \frac{\sum\limits_{u\in U}
r_{u,A}r_{u,B}}{\sqrt{\sum\limits_{u\in U} r_{u,A}^2}
\sqrt{\sum\limits_{u\in U} r_{u,B}^2}},
\end{align}
where $A$ and $B$ are items, $U$ is the set of all users and $r_{u,I}$ is the 
rating given by user $u$ to item $I$.

In the second model, we imagine an unlimited supply of users, each of whom
chooses to rate a subset of items. The lack of a rating is now meaningful since
it indicates a lack of interest of a particular user for a particular item. This
is captured by the model through a default rating of zero\footnote{We address
the issue of rating bias and the precise meaning of a zero valued ratings later
in this report.}. At any point in time, we may estimate the true similarity in
this model of {\em preferential rating}, by computing the cosine similarity of
the two item vectors. Only in the limit of infinite time does this estimate
converge to the true similarity, which we denote $TrueSimPref$.  That is,
\begin{align}
TrueSimPref(A, B) = lim_{t\to\infty}\frac{\sum\limits_{u\in U_{AB}}
r_{u,A}r_{u,B}}{\sqrt{\sum\limits_{u\in U_A} r_{u,A}^2}
\sqrt{\sum\limits_{u\in U_B} r_{u,B}^2}},
\end{align}
where $U_A$ is the set of users who have rated $A$, $U_B$ is the set of users
who have rated $B$, and $U_{AB}$ is the set of users who have rated both $A$ and
$B$.

\section*{Estimating Item-Item Similarity}

Of course, most systems will never acquire ratings from all users for all items.
Nor are we able to work in the limit of infinite time. We must therefore make
estimates of the true item-item similarity using only partial information. In
the case of $TrueSimPref$, the cosine similarity computed at any point in time
serves as a simple and obvious estimate. In using this estimate, we are treating
all missing ratings as intentionally omitted. When estimating $TrueSimRand$, the
choice of how to handle missing ratings is perhaps less obvious. Following the
convention of Breese et al. \cite{Breese1998}, we choose to use the Pearson
correlation\footnote{The Pearson correlation is traditionally characterized by
subtracting an expectation value from each dimension. This is only approximately
true in our case. We discuss how we subtract rating biases later in this
report.} for this estimate, ignoring all ratings by users who have not rated
both items. To summarize, our naive estimates are
\begin{align}
PearSim(A, B) = \frac{\sum\limits_{u\in U_{AB}}
r_{u,A}r_{u,B}}{\sqrt{\sum\limits_{u\in U_{AB}} r_{u,A}^2}
\sqrt{\sum\limits_{u\in U_{AB}} r_{u,B}^2}}
\stackrel{?}{\approx} TrueSimRand(A, B)
\end{align}

\begin{align}
CosSim(A, B) = \frac{\sum\limits_{u\in U_{AB}}
r_{u,A}r_{u,B}}{\sqrt{\sum\limits_{u\in U_A} r_{u,A}^2}
\sqrt{\sum\limits_{u\in U_B} r_{u,B}^2}}
\stackrel{?}{\approx} TrueSimPref(A, B).
\end{align}

\section*{Problem}
Both similarity estimates primarily leverage information from common users, that
is users who have rated both items. Because of this, they perform well with a
large number of common users, but give unreliable, and dramatically varying,
results when the items have few common users. In an extreme case, if only one
user has rated item $A$ and item $B$, $PearsSim(A, B) = 1$ or $-1$. Not only is
this unlikely to be an accurate assessment of the true similarity of $A$ and
$B$, it also gives the most extreme result possible, without giving any
indication that this is a low-confidence calculation.

In this paper, we present two approaches to better approximate the
true similarity given a limited number of common reviewers. Both approaches are
motivated by probabilistic models with distributions grounded in the data. The
first method leverages the observed naive similarity estimate and the number of
common users to better estimate the true similarity. The second method leverages
the observed ratings themselves for the same ends.

\subsection*{Data}
In the following section, we will describe an off-line experiment used to
evaluate our similarity estimators. But first, we provide some details about the
data used for this work. The data, originally from Amazon.com, was acquired from
the Stanford Network Analysis Project\footnote{http://snap.stanford.edu}. It
consists of ~5.6M ratings by ~1.2M users of ~600k music items (e.g. albums and
songs). Each rating is from 1 to 5 stars and is associated with a particular
user and item at a particular time.  For practical reasons, we focus our work on
two subsets of this data. The first consists of the 13,752 item pairs having at
least 40 common users (i.e. users that have provided a rating for both items).
The second consists of the 1,312 item pairs having at least 80 common reviewers.
As discussed below, we must confine ourselves to item pairs that have many
common reviewers in order to determine a good approximation of their true
similarity.

Before any analysis of the data, we remove the biases from each rating, as in
\cite{Koren2009}:
\begin{align}
\hat{r_{ui}} = r_{ui} - \mu_r - b_u - b_i
\end{align}
where $r_{ui}$ is the raw rating user $u$ gave item $i$, $\mu_r$ is the average
rating over the entire dataset, $b_u$ is the user bias, calculated as the
average rating from user $u$ after $\mu_r$ has been subtracted from each, and
$b_i$ is the item bias, calculated as the average rating of item $i$ after
$\mu_r$ and $b_u$ have been subtracted from each.

\subsection*{Experiment}
Ideally, we would know {\em a priori} the true similarity for a set of item
pairs. We could then generate similarity estimates and readily determine the
true error in these estimates. But both $TrueSimRand$ and $TrueSimPref$ are
theoretical constructs and not metrics we have access to in the real world.
Fortunately, we can achieve good approximations of either $TrueSim$ given
sufficient data. We therefore define $GoldSimRand = PearSim_N$ and $GoldSimPref =
CosSim_N$ where $N$ is a large enough number of common users that $GoldSim 
\approx TrueSim$.  With a measurable $GoldSim$ as our goal, we can implement 
a simple predictivity experiment.

To obtain good $GoldSim$ values, we work only with the item pairs that have at
least either $N=40$ or $80$ common users. These choices are a compromise between
having enough data for statistically significant results and having a
sufficiently large $N$ that $GoldSim \approx TrueSim$.

In order to evaluate estimates of $TrueSim$, we construct an off-line experiment
to simulate observing sparse amounts of rating data. Our experiment reenacts the
passing of time. Starting with all ratings withheld, we apply ratings to items
in the order they actually occured. The similarity of item pairs is then
estimated each time the number of common users $n$ is incremented. In this way,
we can quantitatively compare various estimators by calculating their mean
squared error. Since different estimators may perform better for different
degrees of sparsity, we compute the mean squared error for $n=1,...,N$
separately.

\section*{Score-based Method}
\subsection*{Probabilistic Model}

In our first approach, we model the probability distributions of true and
observed similarity. The goal is to use the observed similarity score directly
to approximate the true similarity. In this way, we can consider an abstact
$TrueSim$ which ignores the details of $TrueSimRand$ or $TrueSimPref$.
Similarly, we consider $Sim_n$ to be a generic observed similarity score when
the two items have $n$ common reviewers. Let $Y$ be a random variable
representing the $TrueSim$ of a randomly chosen pair of items. Let $X_{n}$ be a
random variable representing the $Sim_{n}$ of a pair of items when $n$ common
users are observed.

We model $Y$ as a Normal distribution 
\begin{align}
Y \sim N(\mu_s, \sigma_{s}^2),
\end{align}
where $\mu_s$ and $\sigma_{s}^2$ are the average similarity score and the variance
of similarity scores of all pairs of items, respectively. Figure 1 shows the
distributions of $GoldSim$, which motivates the Gaussian model for $TrueSim$.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.8\textwidth]{Histograms.png}
	\caption{Histograms of {\em GoldSimRand} (top) and {\em GoldSimPref}
(bottom) for training datasets containing at least 40 common reviewers (left)
and 80 common reviewers (right).}
    \label{fig:Histograms}
\end{figure}

Since $X_n$ represents
a noisy reading of the true similarity $Y$, we model $X_n | Y=y$ as a Gaussian
error around $y$:
\begin{align}
(X_n | Y=y) \sim N(y, \sigma_{err, n}^2)
\end{align}
Note that $\sigma_{err, n}^2$ represents how noisy $Sim_n$ is as an estimate of
$TrueSim$ and therefore should decrease as $n$ increases. The probability
distribution of $X_n$ is illustrated in Figure 2.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.7\textwidth]{twonormals.png}
	\caption{The blue curve represents $Y$, the distribution of true
    similarities between a random pair of items. The red curve represents $X|Y$,
    the observed similarity when a small number of common users exist.}
    \label{fig:two_normals}
\end{figure}

The problem of approximating $TrueSim$ given $Sim_n$ can now be seen as finding
the value of $Y$ that most likely produced $X_n$. In other words we desire the
Maximum Likelihood Estimate of $Y | X_n$:
\begin{align}
\hat{y} &= \argmax_yP(Y=y|X_n=x) 
\\&= \argmax_y\left[\frac{P(X_n=x | Y=y)P(Y=y)}{P(X_n=x)}\right]
\\&= \argmax_y\left[P(X_n=x|Y=y)P(Y=y)\right]
\\&= 
\argmax_y\left[\frac{1}{\sigma_{err,n}\sqrt{2\pi}}\exp{\left(\frac{-(x-y)^2}
{2\sigma_{err,n}^2}\right)}
\frac{1}{\sigma_{s}\sqrt{2\pi}}\exp{\left(\frac{-(y-\mu_s)^2}
{2\sigma_{s}^2}\right)}\right]
\\&= \argmin_y\left[\frac{(x-y)^2}{2\sigma_{err,n}^2} +
\frac{(y-\mu_s)^2}{2\sigma_{s}^2}\right]
\\&= \argmin_y\left[\left(\sigma_{s}^2+\sigma_{err,n}^2\right)y^2 - 
2\left(\sigma_{s}^2x+\sigma_{err,n}^2\mu_s\right)y\right]
\end{align}
We find the exact minimum by taking the derivative with respect to $y$ and
setting it to zero:
\begin{align}
\frac{d}{dy}\left[\left(\sigma_{s}^2+\sigma_{err,n}^2\right)y^2 - 
2\left(\sigma_{s}^2x+\sigma_{err,n}^2\mu_s\right)y\right]
&= 2\left(\sigma_{s}^2+\sigma_{err,n}^2\right)y - 
2\left(\sigma_{s}^2x+\sigma_{err,n}^2\mu_s\right) 
\\&= 0
\end{align}
This produces the following linear equation for $\hat{y}$ in terms of $x$
\begin{align}
\hat{y} &=
\frac{\sigma_{s}^2x+\sigma_{err,n}^2\mu_s}{\sigma_{s}^2+\sigma_{err,n}^2},
\end{align}
which can be rewritten as
\begin{align}
\left(\hat{y} - \mu_s\right) &= \frac{\sigma_{s}^2}{\sigma_{s}^2+\sigma_{err,n}^2}
\left(x-\mu_s\right).
\end{align}

The model suggests that there is a linear correlation between $X_n$ ($Sim_n$)
and $Y$ ($TrueSim$).\footnote{It is worth noting that despite the resemblance to
Slope One Predictors \cite{Lemire1998}, this is a fundamentally different
technique. Here, we have a linear relationship between observed scores and true
scores, whereas the Slope One model seeks a linear relationship between the
ratings a user gives to different items.} Moreover, the best predicted $TrueSim$
is a linear combination of the average true similarity and the observed
similiarity. Since $\sigma_{err,n}^2$ is nonnegative, the slope is always less
than or equal to one. This means that, on average, the observed distance from $\mu_s$
overestimates the true distance from $\mu_s$. This is reasonable, since the prior
distribution suggests that the true similarity is likely to be near $\mu_s$.

This model requires a different parameter $\sigma_{err,n}^2$ for each $n$.  To
gain more generality, we model $\sigma_{err,n}^2$ as a function of $n$ instead
of considering them independent of each other. Recall that $\sigma_{err,n}^2$ is
the noisiness of the observed $Sim_n$ about $TrueSim$.  $Sim_n$ converges to
$TrueSim$ as the number of users $n$ increases. Although $Sim$ is not simply a
sample mean, The Central Limit Theorem motivates the intuition that the variance
of $Sim$ would decrease as the square root of $n$. Thus we model
$\sigma_{err,n}^2$ as
\begin{align}
\sigma_{err,n}^2 = \frac{\alpha}{\sqrt{n}},
\end{align}
yielding the single parameter model
\begin{align}
\hat{y}= \frac{\sigma_{s}^2}{\sigma_{s}^2+\frac{\alpha}{\sqrt{n}}}
\left(x -\mu_s\right) + \mu_s.
\end{align}

\subsection*{Technique and Results}
With a measurable $GoldSim$ as our predictive goal, we can apply supervised 
learning techniques. First, we find pairs of items that have at least $N$ common
users and calculate their $GoldSim$. Next we withold some of the users to 
produce estimated $Sim_n$ for $n=1,...,N$. By doing this over many pairs of 
items, we produce pairs of ($Sim_n$, $TrueSim$) across different values of $n$ 
that can be used for training. Now the parameters $\mu_s$, $\sigma_{s}^2$, and 
$\sigma_{err,n}^2$ can be approximated using the method of moments over the 
training data. The values of $\sigma_{err,n}^2$ are fit to the function 
$\frac{\alpha}{\sqrt{n}}$ and the $\alpha$ that minimizes the sum of squared 
error is computed.

The graph of Fig. 3 compares the results of three linear predictors, with
varying degrees of fidelity to the probabilistic model. The most faithful
predictor uses $\mu_s$, $\sigma_{s}^2$ and $\alpha$ as calculated above to
predict the $TrueSim$ using Eqn (19). The next predictor uses the calculated
values of $\sigma_{err,n}^2$ directly, following Eqn (16) for predictions. The
third predictor simply uses a series of linear regressions, one for each $n$.
This model only assumes that for each $n$, $Sim_n$ and $TrueSim$ are linearly
correlated. The linear regression technique trades the generality of the model
for a better fit on the training data.  As the results show, the trade-off is
worth it.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.7\textwidth]{LinearTechniques.png}
	\caption{Comparison of different linear predictors. For each number of
reviewers in common, $n$, the graphs represent the average squared error of
predicted $GoldSim$ calculated from the observed $Sim_n$. As a baseline, the
technique of guessing the average $GoldSim$ is included. For these experiments,
we use $Sim$ = $PearSim$ and $N \geq 80$ for determining $GoldSim$.}
    \label{fig:Linear}
\end{figure}

\section*{Ratings-based Method}
\subsection*{Probabilistic Model}

We now construct a different probabilistic model based on the pairs of ratings
themselves. This model is readily exploited for making estimates of
$TrueSimRand$ and, to a lesser extent, $TrueSimPref$. In addition, this approach
allows us to infer implied predictions of $TrueSimRand$ from individual users 
each time they rate two items. This latter feature allows us to define a new 
metric for users that we call {\em user-predictivity}. User-predictivity allows
us to ascribe varying levels of confidence to a user's ratings and thus modify
the contribution a user makes to a similarity estimate accordingly.

Let $Y$ be a random variable representing the $TrueSimRand$ of a randomly chosen
pair of items. As before, we model $Y$ as a Normal distribution
\begin{align}
Y \sim N(\mu_s, \sigma_s^2),
\end{align}
where $\mu_s$ and $\sigma_s^2$ are the average and variance of all true
similarity scores, respectively. Let $\vec{R} = \left(R_A~~R_B\right)$ be a
random two element vector representing the ratings given to items $A$ and $B$.
We model the conditional distribution of $\vec{R} | Y$ as a multivariate Normal
distribution
\begin{align}
(\vec{R} | Y=y) \sim N(\vec{\mu_r}, \mathbf{\Sigma_y}),
\end{align}
where $\vec{\mu_r} = (\mu_r~~\mu_r)$ are the average ratings and
$\mathbf{\Sigma_y}$ is the $y$-dependent $2\times2$ covariance matrix.  Since
$y$ is the $TrueSimRand$ score, it can be thought of as the Pearson correlation
over the distribution of ratings $R_A | Y$ and $R_B | Y$. Assuming that the
variance of individual ratings $\sigma_r^2$ is independent of $y$, we have 
\begin{align}
y = \frac{\Cov{\left[R_A | Y=y , R_B | Y=y\right]}}{\sigma_r^2},
\end{align}
It therefore follows that if the ratings are indeed normally distributed 
then the covariance matrix is related to the true similarity $y$ by
\begin{align}
\mathbf{\Sigma_y} = \sigma_r^2 \cdot
\left[ \begin{array}{cc}  1 & y \\ y & 1 \end{array} \right].
\end{align}

Using this model, we may view the problem of estimating the similarity $Y$ of
items $A$ and $B$ that most likely produced ratings $R_A$ and $R_B$ as
that of finding the Maximum Likelihood Estimate of $Y | \vec{R}$:
\begin{align}
\hat{y} &= \argmax_yP(Y=y|\vec{R}=\vec{r}) 
\\&= \argmax_y\left[\frac{P(\vec{R}=\vec{r} |
Y=y)P(Y=y)}{P(\vec{R}=\vec{r})}\right]
\\&= \argmax_y\left[P(\vec{R}=\vec{r}|Y=y)P(Y=y)\right]
\\&=
\argmax_y\left[\frac{1}{2\pi\sqrt{\left|\mathbf{\Sigma_y}\right|}}\exp{\left(-\frac{1}{2}
(\vec{r} - \vec{\mu_r})^\mathrm{T}\mathbf{\Sigma_y}^{-1}(\vec{r} - \vec{\mu_r})\right)}
\frac{1}{\sigma_{s}\sqrt{2\pi}}\exp{\left(\frac{-(y-\mu_s)^2}
{2\sigma_{s}^2}\right)}\right]
\\&= \argmin_y\left[(\vec{r} -
\vec{\mu_r})^\mathrm{T}\mathbf{\Sigma_y}^{-1}(\vec{r} - \vec{\mu_r}) + \frac{(y -
\mu_s)^2}{\sigma_{s}^2} + \ln{\left|\mathbf{\Sigma_y}\right|} \right]
\\&=
\argmin_y\left[\vec{r^\prime}^\mathrm{T}\mathbf{\Sigma_y}^{-1}\vec{r^\prime}
 + \frac{(y - \mu_s)^2}{\sigma_{s}^2} + \ln{\left|\mathbf{\Sigma_y}\right|} \right]
\\&= \argmin_y\left[\frac{1}{\sigma_r^2}\frac{1}{1 -
y^2}\left[r_A^\prime(r_A^\prime - yr_B^\prime) + r_B^\prime(r_B^\prime -
yr_A^\prime)\right] + \frac{(y - \mu_s)^2}{\sigma_{s}^2} +
\ln{\left[\sigma_r^2(1 - y^2)\right]} \right],
\end{align}
where we used the fact that
\begin{align}
\mathbf{\Sigma_y}^{-1} = \frac{1}{\sigma_r^2}
\left[ \begin{array}{cc}  1 & y \\ y & 1 \end{array} \right]^{-1} =
\frac{1}{\sigma_r^2}\frac{1}{1 - y^2}\left[ \begin{array}{cc}  1 & -y \\ -y & 1
\end{array} \right].
\end{align}
and that
\begin{align}
\left|\mathbf{\Sigma}\right| = \sigma_r^2(1 - y^2)
\end{align}
along with the substitution $\vec{r^\prime} = (r_A^\prime~~r_B^\prime) = \vec{r}
- \vec{\mu_r}$. Next, we find the minimum by taking the derivative with respect
to $y$ and setting it to zero:
\begin{align}
\frac{d}{dy}\left[ \frac{1}{\sigma_r^2}\frac{1}{1 -
y^2}\left[r_A^\prime(r_A^\prime - yr_B^\prime) + r_B^\prime(r_B^\prime -
yr_A^\prime)\right] + \frac{(y - \mu_s)^2}{\sigma_{s}^2} +
\ln{\left[\sigma_r^2(1 - y^2)\right]} \right] &= 0 \\\label{eqn:estimator}
-r_A^\prime r_B^\prime \hat{y}^2 + ({r_A^\prime}^2 + {r_B^\prime}^2)\hat{y} - r_A^\prime
r_B^\prime + \frac{\sigma_r^2}{\sigma_s^2}(1 - \hat{y}^2)^2(\hat{y} - \mu_s) - \sigma_r^2(1
- \hat{y}^2)\hat{y} &= 0.
\end{align}
This final expression implicitly yields the MLE of $y$ as a function of the
ratings $r_A$ and $r_B$. That is, by constraining the parameters $\mu_r$,
$\sigma_r$, $\mu_s$, and $\sigma_s$ to realistic values (e.g. $\sigma_r,
\sigma_s > 0$), this expression defines a surface that can be used to
efficiently estimate $y$ given a pair of ratings\footnote{Note that this
equation neglects the constraint that $-1 \leq y \leq 1$. As a consequence, some
combinations of ratings and parameters can yield MLEs for $y$ that exceed $1$ in
magnitude. This is easily fixed, however, by applying this additional
constraint.}.

\subsection*{User Predictivity}

As already mentioned, the estimator $PearSim$ is unable to say much about the
similarity of two items when only provided ratings from a single common user. In
this case, $PearSim$ yields $\pm1$ (or the indeterminate value $0/0$ in the case
that both ratings are zero) depending on whether the ratings have the same sign
or not. The value of $\hat{y}$ provided by the above model rectifies this by
generating similarity estimates in the range $[-1, +1]$ given a single pair of
ratings. We therefore consider each user who rates a pair of items as implicitly
predicting the similarity of the items. A key discovery of the present work is
that the accuracy of users' predictions in a training set is correlated with the
accuracy in a test set. We exploit this finding by ascribing a quantity we call
{\em user-predictivity} to each user defined as
\begin{align}\label{eqn:predictivity}
p_u = e^{-\epsilon_u/\lambda},
\end{align}
where $\lambda$ is a parameter to be learned and
\begin{align}\label{eqn:userError}
\epsilon_u = \frac{1}{\left|I_u\right|}
\sum_{A,B \in I_u}{\left|\hat{y}(r_{u,A},r_{u,B}) - GoldSim(A, B)\right|}
\end{align}
with $I_u$ the set of item pairs rated by user $u$. We are once again using
$GoldSim$ as a proxy for $TrueSim$ (See {\bf Experiment} section) such that
$\epsilon_u$ is the average absolute error of user $u$'s similarity predictions.

\subsection*{Technique and Results}

The parameter values $\mu_s$ and $\sigma_s$ are obtained by fitting the
distribution of $GoldSim$ values in the training set (see Fig.
\ref{fig:Histograms}). For example, in the $N \geq 80$ dataset we have $\mu_s =
0.27$ and $\sigma_s = 0.25$. The choice of $\mu_r$ and $\sigma_r$ is less clear.
Some insight can be gained by examining the prior distribution of ratings for 
our datasets, which are shown in Fig. \ref{fig:RatingDist}. From there, we
experiment with different values of these parameters to minimize the resulting
error over the training set. We find $\mu_r = 0.25$ and $\sigma_r = 0.3$ to be
good parameters and use them when estimating $TrueSimRand$.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.7\textwidth]{RatingDist_40.png}
    \includegraphics[width=0.7\textwidth]{RatingDist_80.png}
	\caption{Distribution of rating pairs by common users for item pairs.  Top
figure shows distribution for item pairs having at least 40 common users. Bottom
figure shows distribution for item pairs having at least 80 common users.
Rating1 and Rating2 are chosen in arbitrary order.}
    \label{fig:RatingDist}
\end{figure}

Figure \ref{fig:RatingDist} shows density plots of rating pairs, for $N \geq 40$
and $N \geq 80$. These data, together with the similarity score histograms of Fig.
\ref{fig:Histograms}, suggest that the data for our two data sets are similar
and that large biases are not introduced by varying the number of common users,
$N$. At the same time, both rating distributions reveal significant correlations
between Rating1 and Rating2 as evidenced by the elongation in the $45^\circ$
direction. This is consistent with the data of Fig. \ref{fig:Histograms}, which
shows the mean $PearSim > 0$ for both sets. Further, these data indicate a
bias toward positive values among rating pairs compared to those of ratings in
general, which have an average value of zero by construction\footnote{See
discussion of bias subtraction in {\bf Data} section.}. This latter feature
indicates that items with large numbers of common users tend to be rated higher
by their common users than by their other users since we have already subtracted
the item's average rating thereby avoiding any simple popularity effects.

We observe that while the data of Fig. \ref{fig:RatingDist} do not
conform precisely to a multivariate Gaussian distribution, this approximation is
not an unreasonable one. In passing, we remark on the faint clusters that flank
the central cluster in the figures. We understand these to be artifacts of the
bias subtraction. Moreover, a more sophisticated Rating-based model could
account for these by considering a $P(\vec{R} | Y)$ that is a mixture of
Gaussians.

Having chosen values for $\mu_s$, $\sigma_s$, $\mu_r$, and $\sigma_r$ we are now
able to estimate the similarity of item pairs. To do this, we must solve Eq.
(\ref{eqn:estimator}) for $\hat{y}$ given a pair of ratings. This can be easily
done in practice using a root finding algorithm\footnote{To solve Eq.
(\ref{eqn:estimator}) for $\hat{y}$ given ratings $r_A$ and $r_B$, we use the
{\tt brentq} root finding function of the widely available Python module {\tt
scipy.optimize}.}. The implied predictions of individual users can then be
aggregated to form a new estimate of the similarity of items $A$ and $B$ as
\begin{align}\label{eqn:userPredSim}
UserPredSim(A,B) = \frac{\sum_{u \in U_{AB}}{p_u\cdot\hat{y}(r_{u,A}, r_{u,B})}}
{\sum_{u \in U_{AB}}{p_u}},
\end{align}
where, as before, $U_{AB}$ is the set of common users for items $A$ and $B$ and
$p_u$ is the {\em user-predictivity} defined in Eq. (\ref{eqn:predictivity}). We
leverage the training set to find $p_u$ for each user $u$. This is done by
recording the implied prediction $\hat{y}(r_{u,A},r_{u,B})$ for each common user
$u$ for each item pair $(A, B)$ in the training set and then computing each
user's average absolute error $\epsilon_u$ as per Eq. (\ref{eqn:userError}).
Finally, we apply supervised learning techniques to optimize the value of the
parameter $\lambda$ that appears in the definition of $p_u$. This is achieved by
minimizing the error in Eq. {\ref{eqn:userPredSim} made against the training
set. We find that $\lambda = 0.01$, which yields a strong preference for high
{\em user-predictivity}, works well for the data studied here.

Figure \ref{fig:MSE_SimR} shows the results of our experiment using
$UserPredSim$ to estimate $TrueSimRand$. To be clear, these plots are showing
the mean squared error (MSE) given by
\begin{align}
\mathrm{MSE} = \frac{1}{\left|S\right|}\sum_{(A,B) \in S}
{\left(Estimator(A,B) - GoldSim(A,B)\right)^2},
\end{align}
where $S$ is the test set of item pairs and $Estimator(A, B)$ represents an
arbitrary estimator of the similarity between items $A$ and $B$. In the upper
panel, we show the MSE for the $N \geq 40$ data. The black dashed line
represents the estimates of $PearSim$. For example, with $n = 20$ common users,
the MSE of $PearSim$ is about $0.07$, which corresponds to an average absolute
error of $0.26$ in the similarity score---roughly the magnitude of $\sigma_s$
(see Fig. \ref{fig:Histograms}). These data clearly show that the reliability of
$PearSim$ in estimating $TrueSimRand$ deteriorates rapidly as $n$ decreases.
Below about $n = 20$, the horizontal red line, which represents the MSE obtained
from always guessing the average $GoldSimRand$ value from the training set,
outperforms $PearSim$. These two estimators serve as simple benchmarks.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.9\textwidth]{MSE_SimR_40.png}
    \includegraphics[width=0.9\textwidth]{MSE_SimR_80.png}
	\caption{Mean square error {\em vs} number of common users, $n$, for
estimates of {\em TrueSimRand} for item pairs ultimately having at least 40
common users (top) and 80 common users (bottom). The dashed curve
represents {\em PearSim} computed using only the ratings from the first $n$
users. The flat red line indicate the MSE that results from always guessing
the average {\em TrueSimRand} from the training set. The green curve is {\em
LinRegSim} and the blue curve is {\em UserPredSim}. }
    \label{fig:MSE_SimR}
\end{figure}

The green curve in the upper panel of Fig. \ref{fig:MSE_SimR} is the MSE of
$LinRegSim$, which was already shown in Fig. \ref{fig:Linear}, and is included
here for comparison.  $LinRegSim$ outperforms always-guessing-the-average for
all $n > 2$ and does better than $PearSim$ over the entire scope of the plot.
$UserPredSim$ is shown in blue and for $3 < n < 33$ achieves a lower MSE than
$LinRegSim$.

The lower panel of Fig. \ref{fig:MSE_SimR} shows the MSE of estimators in the
experiment carried out with item pairs with $N \geq 80$. The curves for
$PearSim$, always guessing the average, and $LinRegSim$ exhibit behavior
analogous to that observed in the $N \geq 40$ experiment, with addition of
increased noise. The performance of $UserPredSim$, meanwhile, is significantly
improved, achieving the MSE that $PearSim$ reaches at $n \approx 70$ by $n
\approx 20$. This result demonstrates the efficacy, at least in some datasets,
of measuring and exploiting {\em user-predictivity} when estimating item-item
similarity.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.9\textwidth]{MSE_SimP_80.png}
	\caption{Mean square error {\em vs} number of common reviewers, $n$, for
estimates of {\em TrueSimPref} for item pairs ultimately having at least 80
common reviewers. The dashed curve represents {\em CosSim} computed using only
the ratings from the first $n$ reviewers. The flat red line indicate the MSE
that results from always guessing the average {\em TrueSimPref} from the
training set. The green curve is {\em LinRegSim} and the blue curve is {\em
UserPredSim}.  }
    \label{fig:MSE_SimP}
\end{figure}

Next, we consider the results of using $UserPredSim$ to estimate $TrueSimPref$.
Recall that $UserSimPref$ corresponds to $CosSim$ as time goes to infinity. As
such, the relationship in our model between the covariance of the conditional
probability distributions that generate ratings given a similarity does not hold
for $TrueSimPref$. That is, in our model
\begin{align}
\mathbf{\Sigma_z} \neq \sigma_r^2 \cdot
\left[ \begin{array}{cc}  1 & z \\ z & 1 \end{array} \right],
\end{align}
where $\mathbf{\Sigma_z}$ is the covariance matrix of $P(\vec{R} | Z=z)$ with
$Z$ a random variable representing the $TrueSimPref$ of two items. Nonetheless,
we can choose parameters and naively use $UserPredSim$ to estimate
$TrueSimPref$. For this experiment, we choose $\mu_s = 0.054$, $\sigma_s =
0.049$, $\mu_r = 0.25$, $\sigma_r = 0.15$, and $\lambda = 0.002$ using the same
methods as before.

In Fig. \ref{fig:MSE_SimP} we show the results of our experiments estimating
$TrueSimPref$ using the $N \geq 80$ data set. The MSE of $CosSim$, like that of
$PearSim$ in the case of $TrueSimRand$, rapidly diverges as the number of common
users $n$ goes to zero. Interestingly, $CosSim$ outperforms
always-guess-the-average for all $n > 25$, in contrast with $PearSim$, which
does the same only after $n > 45$. The speaks to the greater stability of
$CosSim$, which takes into consideration all ratings, not only those from common
users. Meanwhile, the $LinRegSim$ performs surprisingly similar for estimations
of $TrueSimPref$ and $TrueSimRand$. In both cases, the MSE begins at
approximately that of always-guess-the-average for $n = 1$ and then decreases
approximately linearly with increasing $n$. Indeed, this was also true for the
$N \geq 40$ experiment. We attribute this robustness to the fact that the model
from which $LinRegSim$ is derived abstracts away the underlying mechanism from
which $TrueSim$ is computed. Finally, $UserPredSim$, perhaps surprisingly, has
the lowest MSE for $n < 15$, although not by much. For all $n > 15$, $LinRegSim$
outperforms $UserPredSim$ in estimating $TrueSimPref$. And for all $n > 40$,
$CosSim$ achieves lower MSE than that of $UserPredSim$. This lackluster
performance by $UserPredSim$ might be expected, however, given the model on
which it is based is a dubious fit for $TrueSimPref$.

\section*{Conclusion}

In this report, we explored the issue of item-item similarity in the context of
collaborative filtering. Estimates of item-item similarity such as the Pearson
correlation and cosine similarity can have very large variances as data is
acquired, yielding large discrepancies between early results and those computed
later after sufficient rating data is acquired. We described an off-line
experiment that simulates the accumulation of rating data over time and the
impact that this data has on the performance of various similarity estimators.
Two new similarity estimators were described and analyzed.

The first is based on a model of noisy score observations. Using a series of
linear regressions, this estimator consistently performs as well or better than
both the Pearson correlation or cosine similarity at estimating future values of
these same functions. The second new estimator is based on a model of noisy
rating pairs generated by an item pair having a true, but unknown, similarity.
This method uses a parametrized surface to efficiently estimate item-item
similarity given only a single pair of ratings. Combining these implied
similarity ``predictions'' with the notion of {\em user-predictivity}, we
constructed a novel item-item similarity estimator that performs very well, but
whose effectiveness appears limited to estimations of future Pearson
correlation.

\section*{Future Work}
The results achieved over this dataset were significant. In order to show that
the techniques are general, though, we would like to run the same experiments on
other datasets, with different types of items and user behavior. The 
assumptions made for this model were not specific to the data used, so we 
hope to achieve similar results. It would also be interesting to see how 
different the learned parameters would be for other datasets and how well 
the techniques could perform using the same parameters learned here.

We would also like to incorporate latent factors into our model.
Latent factor techniques, such as those presented in \cite{Koren2009}, have
become very popular since the Netflix Challenge and have proved to be better for
predicting user-item rating than classical CF. One experiment would be to
construct a new $TrueSimLatent$ based on the similarity of two items in 
latent factor space. Then we could apply both of the approaches in this
paper analagously in latent factor space. Another idea is to use latent factors 
as a new input to a model for predicting $TrueSimRand$ or $TrueSimPref$.

One difficulty with each of these experiments is that it is time consuming to 
compute the latent factors at different points in simulated time. Since we could
not recompute the matrix factorization for every new user rating, we would 
instead compute the latent factorization of the whole matrix at a few 
``snap-shots'' in simulated time. We then calculate the latent factors of many
different pairs at this time, each varying in their degrees of certainty. 
Hopefully, this should give enough pairs with dfferent degrees of certainty 
to train and test over.

In a different direction, we believe there are a number of possible applications
of the user-predictivity score presented in this paper. Based on the results
found here, it is reasonable to believe that adding weights to users by their
predictivity would increase the performance of any user-item rating predictor.
The predictivity could also have applications in the problem of Robust
Collaborative Filtering \cite{Mehta2007}, where rating data is particularly noisy
or riddled with spam.

\begin{thebibliography}{9}

\bibitem{Su2009}
    Xiaoyuan Su and Taghi M. Khoshgoftaar.
    ``A Survey of Collaborative Filtering Techniques.''
    \emph{Advances in Artificial Intelligence.} vol. 2009,
    Article ID 421425, 19 pages, 2009. doi:10.1155/2009/421425
\bibitem{Breese1998}
    John S. Breese, David Heckerman, and Carl Kadie, 
    ``Empirical analysis of predictive algorithms for collaborative filtering.'' 
    \emph{Proceedings of the Fourteenth conference on Uncertainty in artificial intelligence.}
    Morgan Kaufmann Publishers Inc., 1998.
\bibitem{Lemire1998}
    Daniel Lemire and Anna Maclachlan.
    ``Slope One Predictors for Online Rating-Based Collaborative Filtering.''
    \emph{SDM.} Vol. 5, 2005.
\bibitem{Koren2009}
    Yehuda Koren, Robert Bell, and Chris Volinsky. 
    ``Matrix factorization techniques for recommender systems.''
    \emph{Computer} 42.8, 2009. 30-37.
\bibitem{Mehta2007}
    Bhaskar Mehta, Thomas Hofmann, and Wolfgang Nejdl.
    ``Robust collaborative filtering.'' 
    \emph{Proceedings of the 2007 ACM conference on Recommender systems.}
    ACM, 2007.
\end{thebibliography}

\end{document}